Network Configuration Synthesis with Abstract Topologies

### 2016-11-15 paper/abstraction/paper.tex
minor edits

\newcommand{\inferencelocal}[2]{
    \node[] at (#1 + 1.2, 2.5 / 2 + 2.5) {\textbf{I-local}};
    \inferencenobox{#1}{L$_m(j,k)$}{}{L$_n(j',k')$}{}{}{}
    \inferencenobox{#1+2 + #2}{$\ldots$ L$_P$ L$_m(j,k)$}{}{$\ldots$ L$_P$ L$_n(j',k')$}{}{}{}
    \node[anchor=east] at (#1 + 2 + #2 - .6, 2.5) {P};
    \draw[dashed] (#1+1 + #2 - .2, -.34) rectangle (#1+3 + #2 + .1, 2.85);
}

\newcommand{\inferenceglobal}[2]{
    \node[] at (#1 + 1.2, 2.5 / 2 + 2.5) {\textbf{I-global}};
    \inferencenobox{#1}{L$_1(j,k)$}{}{L$_2(j',k')$}{}{}{}
    \inferencenobox{#1+2+#2}{$\ldots$ L$_P$ L$_m(j,k)$}{}{$\ldots$ A$_Q$ L$_n(j',k')$}{}{}{}
    \node[anchor=east] at (#1 + 2 + #2 - .6, .4) {P};
    \node[anchor=east] at (#1 + 2 + #2 - .6, 2.5) {Q};
    \draw[dashed] (#1+1+#2 - .2, -.34) rectangle (#1+3+#2 + .1, .8);
    \draw[dashed] (#1+1+#2 - .2, 1.7) rectangle (#1+3+#2 + .1, 2.85);
}

@@

The possibility of network failures only exacerbates the difficulty of constructing correct configurations.
Link failures in networks occur frequently; it is not uncommon for a large network to experience dozens of such failures in any given day~\cite{dc-failure-study}. However, existing tools~\cite{condor,propane} can only reason about about fault-tolerance for concrete topologies. The problem for \sysname is significantly harder because it must verify that all possible concrete instantiations of an abstract topology satisfy the desired fault-tolerance requirements, which requires a fundamentally different approach.

%
We frame satisfying the fault-tolerance policy provided by the user as an analysis problem over the structure of the PG. In particular, in this section we develop an analysis that uses information about the structure of  concrete instantiations of the abstract topology to infer bounds on the number of edge-disjoint paths between pairs of concrete nodes.

@@

To learn information about disjoint paths, we develop a collection of sound inference rules for invariants of the above form. Figure~\ref{fig:inference-rules} shows each of the inference rules. The rules are read from bottom to top. The label on the bottom left represents a known invariant. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to new invariants learned after applying the inference rule. Conditions that must be \emph{valid} for the rule to apply are shown in a box next to the rule.

In general, the inference rules try to learn information about both the largest number of disjoint paths to any single node in an abstract role as well as the largest reachable group in a particular role with at least one disjoint path to each.

@@

Finally, rules \infrule{I-local} and \infrule{I-global} allow us to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if there is an inference from $L_m(j,k)$ to $L_n(j',k')$ on the left, then if both nodes are in the same local pod P, we can leave all labels up to P unchanged. The global rule \infrule{I-global} says that when the edge goes between pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly applies every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. Inference rule conditions are checked for validity over all concrete topologies using an SMT solver. Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Invariants learned with $j=0$ or $k=0$ provide no useful information and are discarded.

@@

The algorithm was able to infer that any node in the spine role has at least 1 disjoint path and any node in the TL node has at least 2 disjoint paths to a router in role TG. Note that, in this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint. The bound on the fault tolerance between ToR and spine routers can also be used to determine that an aggregation-induced black hole may arise at the border of the datacenter after a single failure in some concrete instantiation.

@@

The first idea is to tag every protocol message with a BGP community that encodes the state of each DFA when advertising a route to a neighbor. Routers then allow route advertisements that correspond to an edge in the PG and block those that do not.
For example, in the concrete PG in figure~\ref{fig:example-compilation}, router S1 will allow an advertisement from its neighbors N1 and N2. If it uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ before sending this to its neighbors. If it uses neighbor N2, then it will add the tag $(0,1)$ instead.

\para{Preference Search}
The second idea is to have each router totally order advertisements from neighbors. For example, to satisfy the policy that leaving through N1 is preferred to leaving through N2, router S1 should prefer a message it hears from N1 over N2. Similarly, router A1 should prefer routes tagged with $(1,0)$ over those tagged with $(0,1)$.
%In general, because BGP is distributed, each router does not have a full view of the entire network when choosing which path to use. Because finding a collection of preferences to ensure correct end-to-end behavior for all failures is a hard problem,
Propane introduced a conservative search strategy for determining route preferences. To make this work for abstract topologies, we modify the search based on the following observation about the structure of the PG:

@@

If we determine that each router has a total order with respect to the simulation relation ($\leq$) for its corresponding PG nodes, then we can simply sort the nodes and prefer advertisements from peers of node $m$ over those of node $m'$ whenever $m' \leq m$. This condition ensures that BGP devices will converge to the most preferred paths in the network even when failures occur.
%
For example, in the concrete PG the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the PG can always match transitions made on the right hand side with respect to the $\geq_{rank}$ relation. This relation does not hold the other way around since $(S1,0,1) \ngeq_{rank} (S1,1,0)$. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received at $S1$ in state $(0,1)$.
%
Notice that in the abstract PG $(S,0,1) \leq (S,1,0)$ also holds.

@@

\para{Concretization}

One observation is that the spine template configuration and concrete spine configurations are structured similarly. We formalize this idea by defining a simple, vendor-neutral language for BGP configurations called \mbgp and formally define compilation for policies over both concrete and abstract topologies to \mbgp. In addition we define two concretization ($con$) functions that translate an abstract \sysname policy and \mbgp configuration to a concrete \sysname policy and concrete \mbgp configuration respectively. Concretization takes a context $\Gamma : \mathit{Var} \rightarrow 2^{\mathit{Prefix} \times V}$ that maps each template variable to a set of pairs of a concrete prefix and topology location where the prefix is owned. Intuitively, the spine template is concretized by replacing each abstract role with the set of all concrete routers that map to that role and each prefix template variable with a new entry in the configuration for each individual prefix provided by the context $\Gamma$.
The main theoretical result is that the compilation and concretization functions commute:

@@

  \vspace{1em}
  \noindent
  $\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$
  \label{thm:concretization}
\end{thm}
%
\vspace{-.8em}
Full definitions of compilation and concretization, as well as the proof of Theorem~\ref{thm:concretization} are included in the appendix.

@@

Suppose that an operator want to expand the concrete datacenter by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

However, template variables tell us exactly where this kind of change can happen. The only way in which the spine routers depend on anything non-local is in the match on prefix template variables. Therefore, we can avoid having to modify the configurations for the spine routers by associating a new unique community tag for template variable {\small \$GP} and then replacing each match on {\small \$GP} with a test on the new community tag. The new ToR must now simply attach the new community tag to advertisements for its prefixes so downstream routers can match against it. Figure~\ref{fig:bgp-configs} (right) shows the templates after the transformation.

>>>>

\newcommand{\inferencelocal}[2]{
    \node[] at (#1 + 1.2, 2.5 / 2 + 2.5) {\textbf{I-local}};
    \inferencenobox{#1}{X}{}{Y}{}{}{}
    \inferencenobox{#1+2 + #2}{$\ldots$ L$_P$ X}{}{$\ldots$ L$_P$ Y}{}{}{}
    \node[anchor=east] at (#1 + 2 + #2 - .6, 2.5) {P};
    \draw[dashed] (#1+1 + #2 - .2, -.34) rectangle (#1+3 + #2 + .1, 2.85);
}

\newcommand{\inferenceglobal}[2]{
    \node[] at (#1 + 1.2, 2.5 / 2 + 2.5) {\textbf{I-global}};
    \inferencenobox{#1}{X}{}{Y}{}{}{}
    \inferencenobox{#1+2+#2}{L$_P$ X}{}{A$_Q$ Y}{}{}{}
    \node[anchor=east] at (#1 + 2 + #2 - .6, .4) {P};
    \node[anchor=east] at (#1 + 2 + #2 - .6, 2.5) {Q};
    \draw[dashed] (#1+1+#2 - .2, -.34) rectangle (#1+3+#2 + .1, .8);
    \draw[dashed] (#1+1+#2 - .2, 1.7) rectangle (#1+3+#2 + .1, 2.85);
}

@@

The possibility of network failures exacerbates the difficulty of constructing correct configurations.
Link failures in networks occur frequently; it is not uncommon for a large network to experience dozens of such failures in any given day~\cite{dc-failure-study}. However, existing tools~\cite{condor,propane} can reason about fault-tolerance only for concrete topologies. The problem for \sysname is harder because it must verify that all possible concrete instantiations of an abstract topology satisfy the fault-tolerance requirements, which requires a fundamentally different approach.

%
We frame satisfying the fault-tolerance policy provided by the user as an analysis problem over the structure of the PG. In particular, we develop an analysis that uses information about the structure of  concrete instantiations of the abstract topology to infer bounds on the number of edge-disjoint paths between pairs of concrete nodes.

@@

To learn about disjoint paths, we develop a collection of sound inference rules for invariants of the form above. Figure~\ref{fig:inference-rules} shows each rule. They are read from bottom to top. The label on the bottom left is a known invariant. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to invariants learned after applying the inference rule. The adjacent box shows the conditions that must be \emph{valid} for the rule to apply.

In general, the inference rules try to learn about both the largest number of disjoint paths to any single node in an abstract role as well as the largest reachable group in a particular role with at least one disjoint path to each.

@@

Finally, rules \infrule{I-local} and \infrule{I-global} reason across pod hierarchies. \infrule{I-local} says that if there is an inference from $L_m(j,k)$ to $L_n(j',k')$ on the left, then if both nodes are in the same pod P, we can leave all labels up to P unchanged. \infrule{I-global} says that when the edge goes across pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.

@@

Our analysis starts from a fixed source location $src$ and repeatedly applies every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. Inference rule conditions are checked for validity over all concrete topologies using an SMT solver. Because the inference rules may continue to yield increasingly larger symbolic expressions, we make the following observation to ensure termination: for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Invariants learned with $j=0$ or $k=0$ provide no useful information and are discarded.

@@

The algorithm was able to infer that any spine router has at least 1 disjoint path and any TL router has at least 2 disjoint paths to a TG router. In this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint. The bound on the fault tolerance between ToR and spine routers can also be used to determine that an aggregation-induced black hole may arise at the datacenter boundary after a single failure in some concrete instantiation.

@@

The first idea is to tag every BGP announcement with a community that encodes the state of each DFA when advertising a route to a neighbor. Routers then allow announcements that correspond to an edge in the PG and block others. 
For example, in the concrete PG in figure~\ref{fig:example-compilation}, router S1 will allow an advertisement from its neighbors N1 and N2. If it uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ before sending this to its neighbors. If it uses neighbor N2, then it will add the tag $(0,1)$ instead.

\para{Preference Search}
The second idea is to have each router totally order advertisements from neighbors. For example, to satisfy the policy that leaving through N1 is preferred to leaving through N2, S1 should prefer a message it hears from N1 over N2. Similarly, A1 should prefer routes tagged with $(1,0)$ over those with $(0,1)$.
%In general, because BGP is distributed, each router does not have a full view of the entire network when choosing which path to use. Because finding a collection of preferences to ensure correct end-to-end behavior for all failures is a hard problem,
Propane introduced a conservative search strategy for determining route preferences. To make it work for abstract topologies, we modify the search based on the following observation about the PG structure:

@@

If we determine that each router has a total order with respect to the simulation relation ($\leq$) for its corresponding PG nodes, then we can simply sort the nodes and prefer advertisements from peers of node $m$ over those of $m'$ whenever $m' \leq m$. This condition ensures that BGP devices will converge to the most preferred paths in the network even when failures occur.
%
For example, in the concrete PG the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the PG can always match transitions made on the right hand side with respect to the $\geq_{rank}$ relation. This relation does not hold the other way around since $(S1,0,1) \ngeq_{rank} (S1,1,0)$. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received at $S1$ in state $(0,1)$.
%
Notice that in the abstract PG $(S,0,1) \leq (S,1,0)$ also holds.

@@

\para{Concretization}
Observe that the spine template and concrete configurations are structured similarly. We formalize this observation by defining a simple, vendor-neutral language for BGP configurations called \mbgp and formally define compilation for policies over both concrete and abstract topologies to \mbgp. In addition we define two concretization ($con$) functions that translate an abstract \sysname policy and \mbgp configuration to a concrete \sysname policy and concrete \mbgp configuration respectively. Concretization takes a context $\Gamma : \mathit{Var} \rightarrow 2^{\mathit{Prefix} \times V}$ that maps each template variable to a set of pairs of a concrete prefix and topology location where the prefix is owned. Intuitively, the spine template is concretized by replacing each abstract role with the set of all concrete routers that map to that role and each prefix template variable with a new entry in the configuration for each individual prefix provided by the context $\Gamma$.
Our main theoretical result is that the compilation and concretization functions commute:

@@

  \vspace{1em}
  \noindent
  $\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$
  \label{thm:concretization}
\end{thm}
%
\vspace{-.8em}
Full definitions of compilation and concretization, as well as the proof of Theorem~\ref{thm:concretization} are included in the appendix. These results also help assign semantics to templates.

@@

Suppose an operator wants to expand the concrete datacenter by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

However, template variables tell us exactly where this kind of change can happen. The only way in which the spine routers depend on anything non-local is in the match on prefix template variables. Therefore, we can prevent modifications to the spine routers' configurations by associating a new unique community tag for template variable {\small \$GP} and then replacing each match on {\small \$GP} with a test on the new community tag. The new ToR must now simply attach the new community tag to advertisements for its prefixes so downstream routers can match against it. Figure~\ref{fig:bgp-configs} (right) shows the templates after the transformation.
